---
title: Land Cover of Canada
author: Pierre vernier
date: 2026-01-09
---

## Description

*From the website*: Land cover information is necessary for a large range of environmental applications related to climate impacts and adaption, emergency response, wildlife habitat, etc. The Canada Centre for Remote Sensing (CCRS) has generated a 30 m land cover map of Canada in response to this need. This land cover dataset is also the Canadian contribution to the 30 m spatial resolution Land Cover Map of North America under a collaboration called the North American Land Change Monitoring System (NALCMS). This land cover dataset for Canada is produced using observation from Operational Land Imager (OLI) Landsat sensor.

- [Go to website](https://radiantearth.github.io/stac-browser/#/external/datacube.services.geo.ca/stac/api/collections/landcover)

## Using R to get the data

### Select area of interest

We will use an area of interest (AOI) that spans the BC-Yukon border.

```{r}
#| message: FALSE
#| warning: FALSE
library(sf)
library(tmap)
library(terra)
bnd <- st_read("D:/caribou/data/little_rancheria_yt/caribou.gpkg", "aoi9999", quiet=TRUE)
bnd3578 <- bnd |> st_transform(3578) # Yukon Albers
bnd3979 <- bnd |> st_transform(3979) |> # Canada Atlas Lambert
  st_buffer(5000)
tm_shape(bnd3578) + 
  tm_borders(col="black", lwd=3, lty="dashed") +
  tm_basemap("Esri.WorldTopoMap")
```

### Download, crop, and save

Now that we have an AOI object, we can download, crop, and save the landcover data. We will do this for all three available maps: 2010, 2015, and 2020. This step only needs to be performed once, which is why we have placed the code within an if statement. If download is set to 1, the code will be run.

```{r}
#| message: FALSE
#| warning: FALSE
download <- 0 # 1 = download, 0 = do not download
if (download==1) {

  # Dowload data
  lc2010 <- "/vsicurl/https://datacube-prod-data-public.s3.ca-central-1.amazonaws.com/store/land/landcover/landcover-2010-classification.tif"
  lc2015 <- "/vsicurl/https://datacube-prod-data-public.s3.ca-central-1.amazonaws.com/store/land/landcover/landcover-2015-classification.tif"
  lc2020 <- "/vsicurl/https://datacube-prod-data-public.s3.ca-central-1.amazonaws.com/store/land/landcover/landcover-2020-classification.tif"
  r2010 <- rast(lc2010)
  r2015 <- rast(lc2015)
  r2020 <- rast(lc2020)

  # Crop rasters
  rc2010 <- crop(r2010, st_bbox(bnd3979), mask = FALSE) |> 
    as.factor() |> # necessary to save color table and ensure proper projection
    project("EPSG:3578") # default: nearest-neighbour

  rc2015 <- crop(r2015, st_bbox(bnd3979), mask = FALSE) |> 
    as.factor() |> # necessary to save color table and ensure proper projection
    project("EPSG:3578") # default: nearest-neighbour

  rc2020 <- crop(r2020, st_bbox(bnd3979), mask = FALSE) |> 
    as.factor() |> # necessary to save color table and ensure proper projection
    project("EPSG:3578") # default: nearest-neighbour

  # Save rasters
  writeRaster(rc2010, "tmp/lcc2010.tif", overwrite = TRUE)
  writeRaster(rc2015, "tmp/lcc2015.tif", overwrite = TRUE)
  writeRaster(rc2020, "tmp/lcc2020.tif", overwrite = TRUE)
}
```

### Plot the results

Finally, we can check the results with a simple plot of the three landcover maps for our AOI>

```{r}
#| message: FALSE
#| warning: FALSE
#| fig-width: 12
if (download==0) {
  rc2010 <- rast("tmp/lcc2010.tif")
  rc2015 <- rast("tmp/lcc2015.tif")
  rc2020 <- rast("tmp/lcc2020.tif")
}

rc <- c(rc2010, rc2015, rc2020)

# View the results
tm_shape(rc) +
  tm_raster() +
  tm_legend_hide() +
  tm_shape(bnd3578) + 
  tm_borders(col="white", lwd=3, lty="dashed")
```
