---
title: Medium resolution DEM - 30 meters
author: Pierre vernier
date: 2026-01-15
---

## Introduction

Medium resolution digital elevation model - 30 meters (MRDEM-30)

> Source: "https://radiantearth.github.io/stac-browser/#/external/datacube.services.geo.ca/stac/api/collections/mrdem-30"

## Using QGIS

See [tutorial prepared by Arnold](https://github.com/beaconsproject/beacons_processing_provider/blob/main/tutorial/mrdem_tutorial.md).

## Using R

Key points:

- The /vsicurl/ prefix enables reading remote files via HTTP/HTTPS without downloading the entire file first
- crop() trims to the bounding box, while mask = TRUE clips to the exact boundary shape
- COGs work particularly well with this approach because they support partial reading, so only the necessary data is downloaded

## Workflow

In this workflow, we are going to download the national MRDEM and clip it to an area of interest (AOI).

**Step 1** - Load libraries and read aoi boundary layer

```{r}
#| message: FALSE
#| warning: FALSE
library(sf)
library(terra)
bnd <- st_read("D:/caribou/data/little_rancheria_yt/caribou.gpkg", "aoi9999", quiet=TRUE) |>
  st_transform(3979) # Canada Atlas Lambert to match mrdem
bnd10 <- bnd |> st_buffer(10000) # buffer by 10-km
```

**Step 2** - Download 

```{r}
#| message: FALSE
#| warning: FALSE
cog <- "/vsicurl/https://canelevation-dem.s3.ca-central-1.amazonaws.com/mrdem-30/mrdem-30-dtm.tif"
rcog <- rast(cog)
plot(rcog)
```

**Step 3** - Crop mrdem to aoi

```{r}
#| message: FALSE
#| warning: FALSE
rc <- crop(rcog, st_bbox(bnd10), mask = FALSE) 
```

**Step 4** - View the results

```{r}
#| message: FALSE
#| warning: FALSE
plot(rc)
plot(vect(bnd), lwd=3, add=TRUE)
```

**Step 5** - Save mrdem

```{r}
#| message: FALSE
#| warning: FALSE
writeRaster(rc, "tmp/mrdem.tif", overwrite = TRUE)
```
